Lifting CDCL to Numerical Domains

### 2016-10-09 algorithm.tex
static analysis equation

@@

\caption{Abstract Conflict Driven Clause Learning $\{ACDCL(S)\}$ \label{Alg:acdcl}}

@@

@@

Given a \rmcmt{formula (F)} and an abstract valuation $v = \top$,  
ACDCL returns {\em safe} if $F$ is unsatisfiable and P is safe. Else, 
ACDCL returns a model that satisfies $F$ and P is {\em unsafe}.  The 
procedure $deduce(P)$ is similar to BCP step in SAT solvers where 
it computes least fix-point with strongest post-condition using 
forward analysis. If the result of $deduce(P)$ 
is BOTTOM ($\perp$), the algorithm terminates with {\em safe}. Else, the 
{\em gamma completeness} check~\cite{sas01} is performed to determine if it is a 
real counterexample. If the {\em gamma completeness} check is successful, then the 
counterexample is real. Else, the algorithm enters into the while($true$)
loop and heuristically picks a meet irreducible for decision. For example, assuming 
interval domain, decisions restrict the range of intervals for variables, so the analysis jumps under a 
greatest fixed-point. Note that the widening operation in abstract interpretation 
jumps above a least fixed-point, so decisions can be viewed as 
dual widening. The procedure $deduce(P)$ is called next to deduce new facts for 
current decision. The algorithm terminates with {\em unsafe} if the result of $deduce(P)$ 
is {\em gamma complete}. Else, the algorithm enters in to $conflict\_analysis()$ phase to learn 
the reason for {\em conflict}. There can be multiple incomparable reasons for conflict --
based on the choice of Unique Implication Point (UIP), ACDCL heuristically choose one. 
A learnt clause must include asserting cuts which guarantees
derivation of new information after backtracking. The clause learning and backtracking continues 
as long as the result of deduction is BOTTOM ($\perp$) or the 
analysis backtracks to decision level 0. If no further backtrack is possible, then the 
algorithm terminates with {\em safe}. Else, the algorithm makes a new decision and 
the above process is repeated until a real counterexample is obtained or the algorithm 
backtracks to decision level 0 after a conflict in which case it returns {\em safe}. 
Currently, ACDCL handles loops in the program by unrolling the bounded loops. 
However, we are developing a new technique for automatic invariant generation using 
ACDCL for unbounded proofs.

A decision is made by calling a procedure $decide(v)$ which 
returns a meet irreducible $d$.  If $d$ refines the 
current abstract value $v$, then the worklist is updated with 
all transformers that are dependant on the current decision 
$d$ along with the update of the corresponding propagation 
and reason trail, $\mathcal{T}$ and $\mathcal{R}$.

>>>>

\subsection{Static Analysis Equations for Safety}
Static program analysis based on abstract interpretation~\cite{DBLP:conf/emsoft/Cousot07} 
perform safety analysis by computing fixed point to infer invariants 
over program variables.  However, bounded model checking (BMC) tries to search 
for a counterexample in a bounded execution trace, by symbolically executing a 
program up to a finite bound.  
Similar to BMC, ACDCL searches for a counterexample by solving the formula shown below.
For a set $N$ of program analysis constraints defined over a set 
of constraint variables $Var = \{X_i| i \in N\}$, representing a program
$\mathcal{P}$ and a concrete domain $\powerset{Env}$, the safety equation 
is given as follows.  Here, $X_{init}, X_{Err}$ denote the initial valuation 
and the \rmcmt{final valuation} of constraint variables.
\[\varphi = X_{init} \subseteq Env \wedge \underset{p,q \in |\mathcal{P}|}
{\bigwedge} \{ X_p \subseteq post_{(p,q)}(X_q) \} \wedge X_{err} \supset \emptyset \] 

@@

\caption{Abstract Conflict Driven Clause Learning $\langle ACDCL(S) \rangle$ \label{Alg:acdcl}}

@@

\subsection{Algorithm for Abstract Conflict Driven Clause Learning}

@@

The algorithm begins with a set $S$ of abstract deduction transformer 
$\widehat{ded_{\varphi}}$ for a formula $\varphi$ over an abstract 
domain $A$, an abstract valuation $\langle v \leftarrow \top \rangle$, 
and an empty propagation trail and reason trail, $\mathcal{T}$, $\mathcal{R}$, 
respectively.  The propagation trail stores all meet irreducibles inferred 
by the abstact model search phase (deductions and decisions), while the 
reason trail maps elements of propagation trail to transformers that were 
used to drive them.

The procedure $deduce(S,v,\mathcal{T},\mathcal{R})$ computes a greatest 
fixed point over the transformers in F that refines the abstract valuation 
$v$, similar to BCP step in SAT solvers.  If the result of $deduce$ 
returns {\em CONFLICT} ($\perp$), the algorithm terminates with {\em safe}. 
Else, the analysis enters into the while($true$) loop and makes a new decision 
by a call to $decide(v)$ which returns a new meet irreducible $d$ that is
consistent to the state of the solver.  If $d$ refines the current abstract 
value $v$, the corresponding propagation and reason trail, $\mathcal{T}$ 
and $\mathcal{R}$ are updated.

For example, a decision in interval domain restrict the range of 
intervals for variables, so the analysis jumps under a 
greatest fixed-point. Note that the widening operation in abstract interpretation 
jumps above a least fixed-point, so decisions can be viewed as 
{\em dual widening}.  The procedure $deduce$ is called next to infer new 
set of meet irreducibles for the current decision.  The model search phase 
alternates between the decision and deduction until the result of deduction 
is either $\gamma$-complete~\cite{gamma} or the deduction leads to a {\em CONFLICT}.  
If $v$ is $\gamma$-complete, then it cannot be refined further.  Thus, the 
algorithm returns the error trace $v$ and terminates with UNSAFE or the 
current abstraction is insufficient to determine the satisfiability of $\varphi$.  
However, in case of {\em CONFLICT}, the algorithm enters in to 
$conflict\_analysis()$ phase to learn the reason for {\em conflict}.  
There can be multiple incomparable reasons for conflict -- based on 
the choice of Unique Implication Point (UIP)~\cite{cdcl}.  ACDCL heuristically 
choose one reason.  A learnt clause must include asserting cuts which 
guarantees derivation of new meet irreducibles after backtracking. The clause 
learning and backtracking continues as long as the result of deduction 
is conflicting ($\perp$), that is, the abstract valuation $v$ does not
abstractly satisfy the formula.  If no further backtrack is possible, 
then the algorithm terminates and $\varphi$ is {\em SAFE}. Else, the 
algorithm makes a new decision and the above process is repeated 
until a real counterexample is obtained or the algorithm backtracks to 
decision level 0 after a conflict in which case it returns {\em SAFE}.

### 2016-10-09 domains.tex
static analysis equation

@@

\begin{definition}{(Meet Irreducibles)} A meet irreducible is a minimum complementable 
element that cannot be expressed in terms of the meet over a set of other 
elements. Formally, a {\em meet irreducible} $M$ in a complete lattice 
structure $A$ is an element $M \in A$ that has the following property.
\[\forall M_1, M_2 \in A, M_1 \meet M2 = M \implies (M = M_1 \lor M = M_2), M \neq \top \]  
\end{definition}

>>>>

\textbf{Program Analysis Constraints:}  A common approach to analyse 
a program $\mathcal{P}$ is to represent $\mathcal{P}$ as set of 
equations called, {\em program analysis constraints}.  The 
program analysis constraints, as shown below, is an assignment, 
or boolean expressions $BExp$ over program variables $Var$ or
a non-deterministic choice (ternary experssion) between statements $s_1$ and
$s_2$.  All the bounded loops are completely unrolled and represented using 
following constraints. Similarly, an assertion is of type $BExp$.  
\[c::= x:= exp | BExp | (exp) ? s_1, s_2\]
For example, figure~\ref{swssa} presents the program analysis constraints for 
a C program in Static Single Assignment (SSA) form.

@@

\begin{definition}{(Meet Irreducibles)} A {\em meet irreducible} $(M)$ 
in a complete lattice structure $A$ is a minimum complementable element 
$M \in A$ that has the following property.
\[\forall M_1, M_2 \in A, M_1 \meet M2 = M \implies (M = M_1 \lor M = M_2), M \neq \top \]  
\end{definition}
